Simple Wind Models

### 2010-06-05 Winds/papers/Theory/First/wind_simple.tex
blah
git-svn-id: https://svn.ucolick.org/xsvn/trunk/radtrans@4985 ac11df5e-a788-c23b-a8e1-c758c323924d

\def\hub{h_{72}^{-1}}
\def\umfp{{\hub \, \rm Mpc}}
\def\mzq{z_q}
\def\zabs{$z_{\rm abs}$}
\def\mzabs{z_{\rm abs}}
\def\msna{{\rm S/N}^{\rm A}_{912}}
\def\sna{S/N$^{\rm A}_{912}$}
\def\mnull{\nu_{\rm 912}}
\def\nnull{$\nu_{\rm 912}$}
\def\intl{\int\limits}
\def\nstatqso{193}
\def\maxoff{0.4}
\def\clls{1.9 \pm 0.2}
\def\alls{5.2 \pm 1.5}
\def\blls{-0.9^{+0.4}_{-0.05}}
\def\cmma{\;\;\; ,}
\def\perd{\;\;\; .}
\def\ltk{\left [ \,}
\def\ltp{\left ( \,}
\def\ltb{\left \{ \,}
\def\rtk{\, \right  ] }
\def\rtp{\, \right  ) }
\def\rtb{\, \right \} }
\def\sci#1{{\; \times \; 10^{#1}}}
\def \rAA {\rm \AA}
\def \zem {$z_{\rm em}$}
\def \mzem {z_{\rm em}}
\def \mzlls {z_{\rm LLS}}
\def \zlls {$z_{\rm LLS}$}
\def \mzend {z_{\rm end}}
\def \zend {$z_{\rm end}$}
\def \mzstrtt {z_{\rm start}^{\rm S/N^A=2}}
\def \zstrtt {$z_{\rm start}^{\rm S/N^A=2}$}
\def \zstrto {$z_{\rm start}^{\rm S/N^A=1}$}
\def \mzstrto {z_{\rm start}^{\rm S/N^A=1}}
\def \zstrt {$z_{\rm start}$}
\def \mzstrt {z_{\rm start}}
\def\smm{\sum\limits}
\def \lll  {$\lambda_{\rm 912}$}
\def \mlll  {\lambda_{\rm 912}}
\def \mtll  {\tau_{\rm 912}}
\def \tll  {$\tau_{\rm 912}$}
\def \mavtll  {<\mtll>}
\def \tigm  {$\tau_{\rm IGM}$}
\def \mtigm  {\tau_{\rm IGM}}
\def \cmm  {cm$^{-2}$}
\def \cmmm {cm$^{-3}$}
\def \kms  {km~s$^{-1}$}
\def \mkms  {{\rm km~s^{-1}}}
\def \lyaf {Ly$\alpha$ forest}
\def \Lya  {Ly$\alpha$}
\def \lya  {Ly$\alpha$}
\def \mlya  {Ly\alpha}
\def \Lyb  {Ly$\beta$}
\def \lyb  {Ly$\beta$}
\def \lyg  {Ly$\gamma$}
\def \ly5  {Ly-5}
\def \ly6  {Ly-6}
\def \ly7  {Ly-7}
\def \nhi  {$N_{\rm HI}$}
\def \mnhi  {N_{\rm HI}}
\def \lnhi {$\log N_{HI}$}
\def \mlnhi {\log N_{HI}}
\def \etal {\textit{et al.}}
\def \ob {$\Omega_b$}
\def \obh {$\Omega_bh^{-2}$}
\def \om {$\Omega_m$}
\def \ol {$\Omega_{\Lambda}$}
\def \gz {$g(z)$}
\def \mgz {g(z)}
\def \lyaf {Lyman--$\alpha$ forest}
\def \fnhi {$f(\mnhi,X)$}
\def \mfnhi {f(\mnhi,X)}
\def \myfnhi {f_{\rm Ly\alpha}(\mnhi,X)}
\def \yfnhi {$f_{\rm Ly\alpha}(\mnhi,X)$}
\def \lfnhi {$f_{\rm LLS}(\mnhi,X)$}
\def \dfnhi {$f_{\rm DLA}(\mnhi,X)$}
\def \sfnhi {$f_{\rm SLLS}(\mnhi,X)$}
\def \mlfnhi {f_{\rm LLS}(\mnhi,X)}
\def \mdfnhi {f_{\rm DLA}(\mnhi,X)}
\def \msfnhi {f_{\rm SLLS}(\mnhi,X)}
\def \ztot {$\Delta z_{\rm TOT}$}
\def \mztot {\Delta z_{\rm TOT}}
\def \mlplls {\ell_{\rm{PLLS}}(z)}
\def \lplls {$\ell_{\rm{PLLS}}(z)$}
\def \mlzlls {\ell_{\rm{LLS}}(z)}
\def \lzlls {$\ell_{\rm{LLS}}(z)$}
\def \mllls {\ell_{\rm{LLS}}(X)}
\def \llls {$\ell_{\rm{LLS}}(X)$}
\def \ldla {$\ell_{\rm{DLA}}(X)$}
\def \lslls{$\ell_{\rm{SLLS}}(X)$}
\def \mlslls{\ell_{\rm{SLLS}}(X)}
\def \nlls {$n_{\rm LLS}$}
\def \slls {$\sigma_{\rm LLS}$}
\def \mnlls {n_{\rm LLS}}
\def \mslls {\sigma_{\rm LLS}}
\def \drlls {$\Delta r_{\rm LLS}$}
\def \mdrlls {\Delta r_{\rm LLS}}
\def \mlmfp {\lambda_{\rm mfp}^{912}}
\def \lmfp {$\lambda_{\rm mfp}^{912}$}
\def \mbplls {\beta_{\rm pLLS}}
\def \bplls {$\beta_{\rm pLLS}$}
\def \btlls {$\beta_{\rm LLS}$}
\def \mbtlls {\beta_{\rm LLS}}
\def \lteff {$\tau_{\rm eff,LL}$}
\def \teff {$\tau_{\rm eff,LL}$}
\def \tlya {$\tau_{\rm eff}^{\rm Ly\alpha}$}
\def \mtlya {\tau_{\rm eff}^{\rm Ly\alpha}}
\def \mteff {\tau_{\rm eff}}
\def \mllteff {\mteff^{912}}
\def \llteff {$\mteff^{912}$}
%\def \llsteff {$\tilde\mtll$}
%\def \mllsteff {\tilde\mtll}
\def \mnmin {\mnhi^{\rm min}}
\def \nmin {$\mnhi^{\rm min}$}
\def \O {${\mathcal O}(N,X)$}
\newcommand{\cm}[1]{\, {\rm cm^{#1}}}
\def \snrlim {SNR$_{lim}$}
\def\mglls {\gamma_{\rm LLS}}
\def\mavgt {<\mtll>}

\title{The SDSS-DR7 Survey for Lyman Limit Systems}

\author{
J. Xavier Prochaska\altaffilmark{1}, 
John M. O'Meara\altaffilmark{2}, 
Gabor Worseck\altaffilmark{1} 
%\& Scott Burles\altaffilmark{3}
}
\altaffiltext{1}{Department of Astronomy and Astrophysics, UCO/Lick Observatory, University of California, 1156 High Street, Santa Cruz, CA 95064}
\altaffiltext{2}{Department of Chemistry and Physics, Saint Michael's College.
One Winooski Park, Colchester, VT 05439}

\begin{abstract}
We present the results from a survey for Lyman limit system absorption in the Sloan Digital Sky Survey, Data release 5.  I like abstracts, they are fun!
\end{abstract}

@@

\section{Introduction}
% JO: Introduce LLS
% JO: Mention previous surveys.
% JO:  Leads into survey size \& bias.
Studies of hydrogen  absorption in the lines of sight towards 
distant quasars have served to both define, and in recent years bring
precision to, our cosmological models for the universe.  
The low density, highly ionized  \lyaf\ lines (a.k.a.\ the intergalactic
medium, IGM), with 
\ion{H}{1} column densities $\mnhi < 10^{17.2} \cm{-2}$,
through their aggregate statistical properties (e.g.\ their
flux power spectrum, mean flux, and column density distributions)
constrain cosmological parameters such as the primordial power spectrum
and the baryonic mass density
\citep[e.g.][]{rau98,cwb+03,mcdonald05,tytler04,fpl+08}.
The high-density, predominantly neutral damped \lya\ systems (DLAs),
with $\mnhi \ge 10^{20.3} \cm{-2}$, trace the gas which forms stars, and
likely represent the progenitors of modern-day galaxies 
\citep[e.g.][]{wlf+95,wgp05,pw09}.

The majority of \lyaf\ lines and the DLAs have, through
analysis of their \lya\ lines, precisely measured \nhi\ values
that permit detailed study of their physical propertiees (e.g.\ metallicity).
For systems with intermediate \nhi\ values ($\approx 10^{18} \cm{-2}$),
however, \lya\ and most of the Lyman series lines lie on the
flat portion of the curve-of-growth and the \nhi\ value is difficult
to cosntrain.
On the other hand, these systems are optically thick to ionizing
radiation and impose a readily identify spectral signature in a quasar
spectrum at the Lyman limit.
These so-called Lyman limit systems (LLS), currently the least-well
studied of \ion{H}{1} absorption systems at high redshift,
are the focus of this manuscript.

Historically, the LLS were among the first class of quasar
absorption line (QAL) systems to be surveyed \citep{tytler84}.
This is because their spectral signature is easy to identify
even with low-resolution, low S/N spectra.  The principal challenge
is that the Lyman limit occurs redward of the atmospheric
cutoff only for systems (hence quasars) with redshifts $z>2.7$.
For lower redshifts, one requires spectrometers on space-bourn
ultraviolet sattelites.
By the mid 1990's, samples of several tens of LLS were generated
spanning redshifts $z 0 - 4$ \citep{ssb89,lzt9X,stengler95,storrie96}.
These results were derived from heterogeneous sets of quasars discovered
from a combination of color-selection, radio detection, and
spectroscopic surveys.    The spectra, too, was acquired with a 
diverse set of instrumentation and therefore varying S/N and spectral
resolution mitigating differing sensitivity to the precise optical
depth at the Lyman limit.  Although the results were not fully
consistent with one another, the general picture that resulted
was a rapidly evolving population of absorption systems reasonably
described by a $(1+z)^{1.5}$ power-law.

Cosmologically, the LLS contribute much if not most of the universe's
opacity to ionizing radiation.  Until recently, the observed incidence
of the LLS provided the only direct means of estimating the mean
free path \lmfp\ at any redshift \citep[e.g.][]{meiksin,madau99,fg08b}.
In a companion paper \citep{pwo09}, we have presented a new
technique to measure the \lmfp\ that circumvents any knowledge
of the LLS.  A more precise census of the LLS will serve as a 
consistency check for this \lmfp\ calculation, but is unlikely to
ever again be a competitive approach.  Instead, the incidence of LLS
can be used in combination with estimates of \lmfp\ to assess the
\nhi\ frequency distribution for gas 
with $\mnhi \approx 10^{16-18} \cm{-2}$, 
a regime that is very difficult to explore by studying individual
absorption systems.   Surveys of the LLS are also likely to 
place tight constraints on $z \sim 3$
cosmological simulations that include radiative transfer.

Physically, the nature of systems that give rise to LLS remains
an open question.  The systems with the largest \nhi\ values
(i.e.\ $\mnhi \ge 10^{19} \cm{-2}$, the so-called the super-LLS and
DLAs) are likely associated with the interstellar medium and
outer regions of high $z$ galaxies.  These high \nhi\ systems, however,
are only a subset of the LLS population.  Unfortunately, a proper
modeling of the LLS almost certainly requires careful modeling of 
radiative transfer which has thus far been beyond the scope of
modern computations.  Indeed, the few studies to date have tended
to severely underestimate the incidence of LLS \citep{katz94,gardnerXX,gnedin}.
In recent simulations of high $z$ galaxy formation, however, 
theorists have placed great attention on `streams' of cold gas that
carry fresh material from the IGM to star-forming galaxies 
\citep{keres99,dekel08}.  These cold streams have relatively large
surface densities ($N_{\rm H} \sim 10^{20} \cm{-2}$) and could therefore
produce Lyman limit absorption provided the material has a
non-negligible neutral fraction.  Consequently, an accurate census
of the LLS with redshift may directly constrain the nature and
prevalence of cold streams in the young universe.

A final, yet perhaps most important, motivation for studying the LLS
is that these systems may dominated the census of metals at all epochs.
The majority of LLS are metal-bearing, showing metal-line transitions of 
common low and high-ions \citep[e.g.][]{p99,pks2000}.
Because the estimated ionization corrections for LLS with 
$\mnhi \approx 10^{18} \cm{-2}$ is large, the observations likely
track only a trace amount of the metals.  [last comment]

[Add shot in the dark(s)?]

In this paper, we survey the homogeneous dataset of quasar
spectra from the Sloan Digital Sky Survey (SDSS), using all 
7 public data releases.  Our observational analysis
aims to produce the most precise measurement of 
the LLS incidence paying careful attention to systematic
biases.  The wavelength coverage and data quality of the SDSS
quasar spectra focusing the survey to $z \approx 3.5$.  Future
work will depend on follow-up obseravtions of well-define
quasar samples at other wavelengths.

The paper is organized as follows.  In $\S$~\ref{sec:def},
we present a set of LLS defintions used throughout the manuscript.
We adopt a cosmology with 
$H_0 = 72 \, h_{72} \, \mkms \, \rm Mpc^{-1}$, 
$\Omega_{\rm m} = 0.3$, and $\Omega_\Lambda = 0.7$
and report proper lengths unless specified.

\section{Lyman Limit System Definitions}
\label{sec:def}

The cross-section of \ion{H}{1} at energies above the Lyman limit
may be approximated by:

\begin{equation}
\sigma_{\rm LL}(\nu \ge \mnull) \approx  
  6.35 \sci{-18} \ltp \frac{\nu}{\mnull} \rtp^{-3} \; \cm{2} \cmma
\end{equation}

\begin{equation}
\mnull = E_{912}/h = c/\mlll \cmma 
\end{equation}

\noindent $E_{912} = 1$\,Ryd.  Specifically, 
$\nu_{\rm LL} = 3.29 \sci{15}$Hz and 
$\mlll = 911.7641$\AA.
This implies a continuum opacity at wavelengths $\lambda \le \mlll$,

\begin{equation}
\mtll(\lambda \le \mlll) \approx \frac{\mnhi}{10^{17.2} \cm{-2}} \ltp 
   \frac{\lambda}{\mlll} \rtp^{-3} \cmma
\end{equation}
where \nhi\ is the \ion{H}{1} column density. 
For a gas `cloud' intersecting a background source with intrinsic flux
$F_{\rm int}(\lambda)$, the observed flux $F_{\rm obs}(\lambda)$
blueward of the Lyman limit is

\begin{equation}
F_{\rm obs}(\lambda \le \mlll) = 
  F_{\rm int}(\lambda) \exp \ltk -\mtll(\lambda) \rtk
\label{eqn:flux}
\end{equation}

In what follows, we will define a `standard' Lyman limit system (LLS) to be
one where $\mtll \ge 2$, i.e.\ $\mnhi \ge 10^{17.5} \cm{-2}$.
This corresponds to greater than 85\%\ attenuation of an 
incident ionizing radiation field. 
By this definition, the class of LLS includes systems with
$10^{20.3} \cm{-2} \le \mnhi \le 10^{19} \cm{-2}$ 
\citep[the so-called super-LLS or sub-DLAs; e.g.][]{opb+07} and systems with
$\mnhi \ge 10^{20.3} \cm{-2}$ \citep[the damped \lya\ systems; e.g.][]{wgp05}.
In a few cases, we will distinguish between these `strong' LLS from 
those with lower \nhi, referring to the latter as LLS.
We also note that our $\mtll \ge 2$ definition for LLS differs
from other works which adopted $\mtll \ge 1$ or $\mtll \ge 1.5$.
Unfortunately, there is no well-established criterion.

Observationally, the absorption of a background source
by a $\mtll \ge 2$ LLS is 
readily apparent, even in low S/N spectra.
We define absorbers with $\mnhi < 10^{17.5} \cm{-2}$ as the 
partial Lyman limit systems (pLLSs).  To survey these systems,
one requires higher quality spectra or an alternate approach to the analysis.

We define the redshift of an LLS as

\begin{equation}
\mzlls \equiv \frac{\mlll^{\rm LLS}}{\mlll} - 1
\end{equation}
where $\mlll^{\rm LLS}$ marks the observed onset of LL absorption.
In practice, this is is often estimated from strong Lyman series lines
(e.g.\ \lya, \lyb) that accompany the Lyman limit opacity.

We define the sub-set of LLS that occur within 3000\kms\
of the emission redshift of the background source as proximate LLS
(PLLS).  We separate the analysis of these systems from the rest
to investigate changes in the incidence of optically thick gas
near high $z$ quasars due to, e.g.\ the quasar's radiation field
and local environment $\S$~\ref{sec:prox}.

Finally, we define the observable \lzlls\ as the average
number of LLS detected per unit redshift at a given redshift.
In the previous literature, this quantity is also expressed
as $n(z)$, $dN/dz$ and $dn/dz$.

\section{SDSS Quasar Sample and Spectroscopy}

One of the primary objectives of the Sloan Digital Sky Survey (SDSS)
was to discover $\sim 100,000$ new quasars across the northern sky
\citep{yaa+00}.  The strategy of the SDSS team to achieve this
ambitious goal was a four-fold process:
(i) obtain deep, multi-band images across a large area of the sky;
(ii) select quasar candidates by demanding a point-like point-spread-function
and imposing color criteria that separate the candidates from the
Galactic stellar locus;
(iii) obtain follow-up spectra for a magnitude-limited sample
with a fiber-fed spectrograph.
The details of target selection and quasar completeness with
redshift is described at length in a series of SDSS papers
\citep{rfn+02}, but see \citep{wp09} for a new and more accurate analysis;  
and
(iv) automatically identify quasars and estimate their redshifts
(\zem) through template fitting to the optical spectroscopy.

Of these steps, the second has the greatest 
impact on a survey for high $z$ Lyman limit systems. 
The key issue for our survey is whether the presence 
of an intervening LLS biases the targeting of the
background quasar for follow-up spectroscopy.  
In effect, a high $z$ LLS severely `reddens' the quasar
at the bluest optical wavelengths of the SDSS imaging.  
With this effect in mind, the SDSS team imposed cuts
on the $(u-g)$ color which better separated the quasar
locus in color pace from the stellar locus.  The net effect,
however, is to bias the spectroscopic follow-up against 
quasar sightlines {\it without} foreground LLS
\citep[][hereafter PWO09]{pwo09}.  Our analysis indicates
an important bias for quasars with $\mzem < 3.6$. 
For this reason, we will limit the statistical analysis
to quasars with $\mzem \ge 3.6$, but we also explore this bias
by considering the incidence of LLS toward quasars with $\mzem = 3.4 - 3.6$.

The quasar spectra analyzed in this paper were taken from the Sloan
Digital Sky Survey, Data Release 7 \citep{sdssdr7}.  
We retrieved
the `best' 1D spectrum for every source flagged as a QSO or HIZ\_QSO.
This totalled 102,418 unique spectra\footnote{  
The spectra were processed through our automated algorithms for
finding absorption-line features and damped \lya\ candidates
\citep{phw05,shf06}.  
Approximately 10 of the spectra failed 
to be processed (primarily because the SDSS-repoted emission redshift
is erroneous) and were removed from any subsequent analysis.}.
The SDSS survey employs a 
fiber-fed, dual-camera spectrometer that provides continuous wavelength coverage
from $\lambda \approx 3800-9200$\AA\ at a spectral resolution of 
FWHM~$\approx 150 \mkms$.
The SDSS team employs a custom, data-reduction pipeline that performs
sky subtraction using emprical measurements from fibers placed
to avoid objects detected in the SDSS images.
The majority of data suffer from excessive sky noise at long
wavelengths ($\lambda > 8000$\AA)
and the instrument throughput and atmospheric absorption
limits the sensitivity at the shortest wavelengths ($\lambda < 4200$\AA).

A survey for \ion{H}{1} Lyman limit absorption in quasar spectra
invovles two principle steps.  First, one must assess
the flux at wavelengths near the Lyman limit
in the quasar's rest-frame, $\lambda \lesssim \mlll(1+\mzem)$.  
We discuss our procedure for this step in the following section.  
The second step is to 
estimate the flux at wavelengths blueward of LLS candidates.
There are several characteristics of the SDSS spectroscopy which
negatively affect this estimate.  A generic concern is the poorer instrument
response at the bluest wavelengths.  At the bluest wavelengths,
many of the spectra exhibit a very low signal-to-noise ratio 
and yield flux estimates consistent with zero, even without an
intervening LLS.
Therefore, we have limited our survey to redshifts $\mzlls \ge 3.3$
corresponding to $\lambda > 3920$\AA.
In practice, we restrict the quasar sample to objects with
$\mzem \ge 3.6$ and perform a search for LLS
at all redshifts, but then only analyze absorption systems with
$\mzlls \ge 3.3$.  We also limit the survey to quasars with
$\mzem \le 5$ because the SDSS spectra of higher redshift objects
are generally too low S/N to permit a robust analysis.
Figure~\ref{fig:snz}a presents histograms of the emission redshifts
for the statistical sample and the signal-to-noise of the
absorbed continuum at the Lyman limit
(\sna; see below for the definition).

Another difficulty with the SDSS spectra at blue wavelengths
is that the two-dimensional spectra of faint sources may be improperly traced.
On occasion, the 1D extractions
include flux from a neighboring object and yield a systematic 
overestimate of the flux.  This effect reduces the estimated opacity
for a LLS.  
Another issue, especially with a fiber-fed spectrometer, is that
the sky model is estimated from nearby fibers that are intentionally
placed on `object-free' regions of the sky.  Although the SDSS
project has worked carefully to mitigate the effects of variable
fiber efficiency and XXX \cite{skysub},  significant misestimates 
of the sky occur.  
We have identified tens of objects where the extracted
flux is significantly negative, indicating an overestimate
in the sky model.  
This may convert a partial LLS (with $\mtll < 1$) into
an LLS.  By a similar token, a proper $\mtll \ge 2$ LLS may
appear as a pLLS if the sky is underestimated.
We proceed under the expectation that this effect
is nearly random, i.e.\ for every underestimate of the sky there
is a corresponding overestimate, but this has not been rigorously
established.  
The SDSS fibers are sufficiently wide
(diameter of $3''$) that they will occasionally include flux
from a projected neighbor.  These coincident objects may be
much fainter than the quasar at redder wavelengths, but they
could contribute all of the flux blueward of a strong LLS
and lead to an underestimate of the LL opacity\footnote{An amusing
(and plausible) systematic effect related to this is contamination
by the light reflected from terestrial satellites crossing the night sky.  
Even a brief `exposure' through the 3$''$ fiber could dominate the
flux at the bluest wavelengths, although this should generally be
mitigated by the fact that the SDSS team aquires 
3 unique exposures per target.}.

\section{The Absorbed Quasar Continuum}
\label{sec:continuum}

Absent any other sources of opacity, one can trivially
estimate \tll\ from
the quasar spectrum through measurements of the flux
both redward and blueward of the observed Lyman limit (Equation~\ref{eqn:flux}).  
In practice, however,
the quasar flux is also attenuated by line opacity from the so-called
\lya\ forest (a.k.a., the intergalactic medium; IGM).
For example, consider a Lyman limit system at $\mzlls = 3.5$
intervening a $\mzem = 4$ quasar.  The LLS attenuates the quasar flux
blueward of $\mlll^{\rm LLS} = 4103$\AA.  At this wavelength,
the quasar spectrum recorded on Earth will also include opacity
from the \lya\ forest at 
$z_{\rm \mlya} = (1+\mzlls)(\mlll/\lambda_{\rm Ly\alpha}) - 1$,
\lyb\ absorption from the IGM at 
$z_{\rm Ly\beta} = (1+\mzlls)(\mlll/\lambda_{\rm Ly\beta}) - 1$, etc.
It is necessary, therefore, to account for these additional sources
of opacity when estimating \tll.

We can express
the observed quasar flux $F_{\rm obs}$ in terms of
the intrinsic flux (just) redward of the Lyman limit $F_{\rm int}$ as

\begin{equation}
F_{\rm obs} (\lambda \gtrsim \mlll) = F_{\rm int} \exp [-\mtigm(\lambda)] \cmma
\end{equation}
where \tigm\ is the effective opacity of the IGM from
Lyman series line-opacity\footnote{In the following, we do not
explicitly derive the 
opacity from metals in the IGM, but these may be considered included in \tigm.}.
The LLS introduces an additional, continuous opacity:

\begin{equation}
F_{\rm obs} (\lambda \le \mlll) = 
  F_{\rm int} \exp [-\mtigm(\lambda) -\mtll(\lambda)] \perd
\end{equation}
A precise estimate of \tll, therefore, requires an estimation
of the absorbed quasar flux not its intrinsic flux.  Conveniently,
this quantity is the observed flux recorded in the spectrum at
$\lambda \gtrsim \mlll$.  
There are still signficant challenges because
the IGM opacity is stochastic on both small (individual Lyman lines)
and modest scales (many 10\AA) 
and 
the intrinsic quasar spectrum (both shape and normalization) varies
from source to source.
We now describe an automated procedure
used to infer the absorbed continuum from each quasar spectrum.

The traditional method of estimating the
quasar continuum is to first identify
regions of unabsorbed quasar flux and then to interpolate a continuum level
between these regions.  For quasars at
high redshift, this method is particularly error prone, because at
wavelengths below \lya\ emission we expect few (and at very
high redshift, none) of the pixels to be free
of absorption from the IGM.   Moreover, this
traditional method frequently requires by-hand modification,
which is time-intensive and subjective to individual biases.
Methods do exist to automatically generate a quasar
continuum from emission-line characteristics \citep[e.g.][]{suzuki+05},
but these are designed to infer the intrinsic quasar spectrum
not the IGM-absorbed continuum.
Our approach is to match a template model of the average absorbed
continuum to each spectrum,
allowing for a large-scale tilt (i.e. a unique underlying power-law slope)
and arbitrary normalization.

Our first step is to derive the templates for the average absorbed
quasar continuum.  Because the line-density of the IGM and therefore \tigm\
increase with redshift, we perform this analysis in small redshift intervals
($\delta z = 0.3$ to 0.6; Figure~\ref{fig:template}). 
For every quasar within the redshift interval,
we shift the spectrum to the quasar rest-frame 
using the SDSS reported \zem\ value.  
Next, we "detilt" the spectrum by removing a power-law shape.
This power-law is determined as follows.  We have
constructed from the SDSS-DR3 dataset an average template spectrum,
archived in XIDL\footnote{http://www.ucolick.org/$\sim$xavier/IDL}
as ``full\_SDSS\_LLS.fits''.
For each individual quasar spectrum,
we sample quasar pixels with wavelengths greater 
than \lya\ emission, divide by the template spectrum,
and measure the spectral slope of the resulting spectrum.

After aligning the spectra in the quasar rest-frame
(nearest pixel), we median-combine the data in each redshift 
interval\footnote{This stack is not optimal for deriving \lya\
forest statistics \citep{dlw08}.}.
The resultant template
represents the median intrinsic quasar continuum modulated by the median
flux decrement of the IGM.  Due to the presence of
Lyman limit systems, the quasar templates will not be useful at
wavelengths near and below the rest-frame Lyman limit;
this portion of the spectrum, however, can be used to 
constrain the mean free path to ionizing radiation \citep{pwo09}.
For wavelengths blueward of 920\AA, therefore,
we set the template to have the value recorded 920\AA.
Conveniently, there are no strong emission features in the
quasar SED at these wavelengths \citep[e.g.][]{telfer02}.
The resultant template spectra are shown in Figure \ref{fig:template}.

>>>>

\newcommand{\naid}{\ion{Na}{I}~$\lambda\lambda 5891, 5897$}
\newcommand{\mgiid}{\ion{Mg}{2}~$\lambda\lambda 2796, 2803$}
\newcommand{\mgiia}{\ion{Mg}{2}~$\lambda 2796$}
\newcommand{\mgiib}{\ion{Mg}{2}~$\lambda 2803$}
\def\hub{h_{72}^{-1}}
\def\umfp{{\hub \, \rm Mpc}}
\def\mzq{z_q}
\def\zabs{$z_{\rm abs}$}
\def\mzabs{z_{\rm abs}}
\def\intl{\int\limits}
\def\cmma{\;\;\; ,}
\def\perd{\;\;\; .}
\def\ltk{\left [ \,}
\def\ltp{\left ( \,}
\def\ltb{\left \{ \,}
\def\rtk{\, \right  ] }
\def\rtp{\, \right  ) }
\def\rtb{\, \right \} }
\def\sci#1{{\; \times \; 10^{#1}}}
\def \rAA {\rm \AA}
\def \zem {$z_{\rm em}$}
\def \mzem {z_{\rm em}}
\def\smm{\sum\limits}
\def \cmm  {cm$^{-2}$}
\def \cmmm {cm$^{-3}$}
\def \kms  {km~s$^{-1}$}
\def \mkms  {{\rm km~s^{-1}}}
\def \lyaf {Ly$\alpha$ forest}
\def \Lya  {Ly$\alpha$}
\def \lya  {Ly$\alpha$}
\def \mlya  {Ly\alpha}
\def \Lyb  {Ly$\beta$}
\def \lyb  {Ly$\beta$}
\def \lyg  {Ly$\gamma$}
\def \ly5  {Ly-5}
\def \ly6  {Ly-6}
\def \ly7  {Ly-7}
\def \nhi  {$N_{\rm HI}$}
\def \mnhi  {N_{\rm HI}}
\def \lnhi {$\log N_{HI}$}
\def \mlnhi {\log N_{HI}}
\def \etal {\textit{et al.}}
\def \lyaf {Lyman--$\alpha$ forest}
\def \mnmin {\mnhi^{\rm min}}
\def \nmin {$\mnhi^{\rm min}$}
\def \O {${\mathcal O}(N,X)$}
\newcommand{\cm}[1]{\, {\rm cm^{#1}}}
\def \snrlim {SNR$_{lim}$}

\title{Simple Wind Models}

\author{
J. Xavier Prochaska\altaffilmark{1}, 
Others
%John M. O'Meara\altaffilmark{2}, 
%Gabor Worseck\altaffilmark{1} 
%\& Scott Burles\altaffilmark{3}
}
\altaffiltext{1}{Department of Astronomy and Astrophysics, UCO/Lick Observatory, University of California, 1156 High Street, Santa Cruz, CA 95064}
%\altaffiltext{2}{Department of Chemistry and Physics, Saint Michael's College.
%One Winooski Park, Colchester, VT 05439}

\begin{abstract}
Blah
\end{abstract}

@@

\section{Introduction}

\begin{itemize}
\item Usage of cool gas outflows to probe galactic-scale winds in
  star-forming galaxies is now a well-established observational
  technique
\item Simple anlaysis focusing on the central wavelength and EW of the
  lines
\item \lya, MgII, NaI and (now) FeII* show emission
\item Despite the wealth of observational diagnostics, little effor
  hast been taken to constrain the wind properties
\item These are some first steps by considering idealized wind
  profiles.  Algorithms can be extended to any wind output
\end{itemize}

\section{Methodology}

\subsection{The Radiative Transitions}

[Neglect collisional excitation and de-excitation.  These would
require $n_e > 10^3 \cm{-3}$.  Non-detection of FeI argues against
this.]

In this paper, we will focus on two sets of radiative transitions
associated with the Fe$^+$ and Mg$^+$ ions.  The choice of these
specific low-ions is motivated by our own observations
\citep{rubin10b} but (as described below) these two ions exhibit
the dominant characteristics of the majoriy of low-ions used to study
cool-gas outflows. [word] Therefore, the majority of our results can
be generalized to other transitions that are freuqenly used.

The Mg$^+$ ion, with a single 3s electron in the ground-state,
exhibits an alkali doublet of transitions that are analagous to the
\lya\ transitions of neutral hydrogen.  In Figure~\ref{fig:energy}
we present the energy level diagram for the Mg$^+$ ion associated with
the commonly observed \mgiid\ doublet.  In non-relativistic quantum
mechanics, the 2p$^6$3p level is said to be split by spin-orbit
coupling.  This doublet marks the only \ion{Mg}{2} electric-dipole transitions
with wavelengths near 2800\AA.  Furthermore, the transtion connecting
the $\rm {}^2P_{3/2}$ and $\rm {}^2P_{1/2}$ is forbidden by several
selection rules.  Therefore, an absorption to either of the upper
states is followed essentially 100$\%$ of the time by a spontaneous decay to the
ground state.

The physics of the \mgiid\ doublet
is analagous [repeating myself] to the physics of \ion{H}{1}
\lya, and also the \naid\ doublet and many other doublets commonly
studied in the ISM and IGM.  The key quantitative differences are in
the wavelengths of the transitions and, for the calculations of interest
in ths paper, their energy separations.  For \ion{H}{1} \lya, the
separation is sufficiently small $\Delta v = c \Delta E / E \approx 1.3
\mkms$ that most radiative transfer treatments ignore the fact that
\lya\ is a pair of transitions.   This is justified by the fact that
most astrophysical processes have turbulent processes that
significnatly exceed this velocity separation and effectively mix the
two transitions.  For \ion{Mg}{2} ($\Delta v \approx 770 \mkms$),  
\ion{Na}{1} ($\Delta v \approx 304 \mkms$), and most other doublets
commonly studies the separations are large and demand that the lines
be treated separately.  In terms of the analysis of outflows, the line
separation is especially relevant when the outflow speed matches the
line separation.

Iron exhibits the most complex set of energy levels of the elements
frequently studied in astrophysics.  The Fe$^+$ ion alone has over XXX
energy levels recorded \citep{iron} and even this is a necessarily
incomplete list.  In this paper, we will focus on 9 of the levels
related to the UV1 mutliplet whose wavelengths lie near 2600\AA\
(Figure~\ref{fig:energy}).  One reason for the complexity of iron is
that the majoriyt of energy levels exhibit fine-structure splitting.
This includes the ground-state which has four fine-structure levels with
excitation energies $T_{\rm ex} = \Delta E / k$ ranging from $\approx
500-1500$\,K.  We label these according to the total angular momentum
of each level $J$, with the ground-state having $J=9/2$.
Transitions between these fine-structure levels are magnetic-dipole
transitions with spontaneous decay times of several hours.  In
principle, these levels can be populated by indirect UV pumping
\citep[e.g][]{sv02,pcb06}, but this would require that the gas lie
close to the source of UV photons (see $\S$~\ref{sec:pump}).  In the
following, we will assume that only the ground-state is populated by
electrons and therefore capable of absorbing photons.

[Comment on UV1 a bit more]

\subsection{The Source}

The majority (all?) of observational studies of cool gas outflows in
absorption have focused on intervening, star-forming galaxies.  The
unobscured emisison of these galaxies is a complex combination of
stars and \ion{H}{2} regions that is then modulated by dust and gas
within its ISM.  In terms of the spectral regions studied in this
paper, most stars show a featureless continuum with a few spectral
types showing significant \ion{Mg}{2} and \ion{Fe}{2} absorption.
[Any emission??]  Furthermore, \ion{H}{2} regions are observed to emit
at the \mgiid\ doublet, primarily due to recombinations in the outer
layers.  It is beyond the scope of this paper to properly model this
stellar absorption and \ion{H}{2} region emission, but the reader
should be aeware that they can complicate the observed spectrum,
independent of the effects of a galactic outflow.
In the following, we will simply assume a flat continuum that will
generally be normalized to unit value.

[What does MgI tell us about winds?  Why no P-Cygni?]
